---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# meteospain

<!-- badges: start -->
[![R-CMD-check](https://github.com/emf-creaf/meteospain/workflows/R-CMD-check/badge.svg)](https://github.com/emf-creaf/meteospain/actions)
<!-- badges: end -->

`meteospain` aims to offer access to different spanish meteorological stations data in an uniform way.

## Installation

`meteospain` is yet in active development. You can install the development version from
[GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("emf-creaf/meteospain")
```

## Services

The following meteorological stations services are available:

  - [AEMET](http://www.aemet.es/en/portada), the spanish State Meteorological Agency.
  - [MeteoCat](https://meteo.cat), the Catalan Meteorology Service.
  - [MeteoGalicia](https://www.meteogalicia.gal/web/inicio.action), the galician meteorological service.
  - [RIA](https://www.juntadeandalucia.es/agriculturaypesca/ifapa/riaweb/web/), the andalucian agroclimatic
  information network.
  - [Meteoclimatic](https://www.meteoclimatic.net/), the spanish non-professional meteorological stations 
  network.


## Examples

Access to the services is done with the `get_meteo_from` function, providing the name of the service and
the options. Each service has a dedicated options function to guide thorugh the especifics of each service:

```{r example}
library(meteospain)

mg_options <- meteogalicia_options(resolution = 'current_day')
get_meteo_from('meteogalicia', mg_options)
```

Stations info can be accessed with `get_stations_info_from` function:

```{r stations_info}
get_stations_info_from('meteogalicia', mg_options)
```

Returned objects are spatial objects (thanks to the [`sf`](https://r-spatial.github.io/sf/) R package), so we
can plot the results directly.

```{r plot_stations}
library(sf)
mg_options <- meteogalicia_options(resolution = 'daily', start_date = as.Date('2021-04-25'))
plot(get_meteo_from('meteogalicia', mg_options))

plot(get_stations_info_from('meteogalicia', mg_options))
```
