---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# meteospain

[![R-CMD-check](https://github.com/emf-creaf/meteospain/actions/workflows/R-CMD-check.yaml/badge.svg?branch=main)](https://github.com/emf-creaf/meteospain/actions/workflows/R-CMD-check.yaml)
[![CRAN-status](https://www.r-pkg.org/badges/version/meteospain)](https://CRAN.R-project.org/package=meteospain)
[![CRAN-RStudio-mirror-downloads](https://cranlogs.r-pkg.org/badges/last-month/meteospain?color=blue)](https://r-pkg.org/pkg/meteospain)

`meteospain` aims to offer access to different Spanish meteorological stations data in an uniform way.

## Installation

`meteospain` is in CRAN, and can be installed as any other package:

``` r
install.packages('meteospain')
```

Also, `meteospain` is in active development. You can install the development version from
[GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("emf-creaf/meteospain", ref = "devel")
```

## Services

The following meteorological stations services are available:

  - [AEMET](https://www.aemet.es/en/portada), the Spanish State Meteorological Agency.
  - [MeteoCat](https://meteo.cat), the Catalan Meteorology Service.
  - [MeteoGalicia](https://www.meteogalicia.gal/), the Galician Meteorological Service.
  - [RIA](https://www.juntadeandalucia.es/agriculturaypesca/ifapa/riaweb/web/), the Andalucian Agroclimatic
  Information Network.
  - [Meteoclimatic](https://www.meteoclimatic.net/), the Spanish non-professional meteorological stations 
  network.


## Examples

Access to the services is done with the `get_meteo_from` function, providing the name of the service and
the options.
Each service has a dedicated `*service*_options()` function to guide through the specifics of each service:

```{r example}
library(meteospain)

mg_options <- meteogalicia_options(resolution = 'current_day')
get_meteo_from('meteogalicia', mg_options)
```

Stations info can be accessed with `get_stations_info_from` function:

```{r stations_info}
get_stations_info_from('meteogalicia', mg_options)
```

Returned objects are spatial objects (using the [`sf`](https://r-spatial.github.io/sf/) R package),
so results can be plotted directly:

```{r plot_stations}
library(sf)
mg_options <- meteogalicia_options(resolution = 'daily', start_date = as.Date('2021-04-25'))
plot(get_meteo_from('meteogalicia', mg_options))

plot(get_stations_info_from('meteogalicia', mg_options))
```

## API keys

Some services, like AEMET or Meteocat, require an *API key* to access the data.
`meteospain` **doesn't** provide any key for those services, see `?services_options` for information
about this.

Once a key has been obtained, we can get the meteo:

```{r aemet_key_use}
get_meteo_from('aemet', aemet_options(api_key = keyring::key_get("aemet")))
```

