---
title: "weatherspain specifications"
author: "Víctor Granda"
date: "4/27/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intro

The idea is to extract the meteorlogical station data download functionality from `meteoland` to this
new package. This will involve also an update in both, the code, and the user interface (more tidy friendly,
sf...).

# Expected available workflows

  1. **Data access**: Programmatic access to APIs and RSS feeds to access the different spanish meteorological
  services, all under the same user interface
  
  1. **Info access**: Programmatic access to APIs and RSS feeds to access services stations info (coordinates,
  elevation... if available)

This workflow will be structured as follows:

# Workflow 1: Data access

## OPTION 1

One function (UI), i.e. `get_meteo_from(service, options, ...)`. The advantage of this is that the user
always use the same interface for all services. `options` argument will provide the specific options for
each service (api, dates, stations...). `options` also will allow selecting between historical, current day
or instant, as well as station info.

```{r option_1_examples, eval=FALSE}
get_meteo_from('aemet', options = list(api = 'xxx', dates = '2020-04-25', stations = 'X234'), ...)
```

This also involves creating helper functions to generate the options list for each service, with sensible
defaults:

```{r option_1_helpers, eval = FALSE}
aemet_options <- function(api, dates = Sys.Date(), stations = NULL, ...) {...}
smc_options <- function(api, dates = Sys.Date(), stations = NULL, ...) {...}
# ...
```

And, of course, helper functions to manage the quirks of each service to use internally in `get_meteo_from`:

```{r option_1_internal_helpers, eval=FALSE}
.get_aemet_data <- function(options, ...) {...}
.get_smc_data <- function(options, ...) {...}
# ...
```

This way opens the possibility of piping access to different services and creation of an sf with all the 
desired stations in one take, always using the same interface.

```{r piping_example, eval=FALSE}
all_catalonia_stations <-
  get_meteo_from('aemet', aemet_options(), ...) %>%
  dplyr::bind_rows(get_meteo_from('smc', smc_options(), ...)) %>%
  dplyr::bind_rows(get_meteo_from('meteoclimatic', meteoclimatic_options(), ...))
```

As a con, internal code for each service (`.get_*_data` helpers) is going to be a little more complicated
to take into account different possibilities (hourly, daily, monthly, current, instant). But this is
unavoidable in any case, so...

## OPTION 2

"meteoland style". Meaning, a function for each service and data type (hourly, daily, monthly...). This also
allows a piping style, but more complex, as each function will have its own interface (some have apis,
some don't, some offers hourly, some don't...). This is not a big concern right now, but if we are adding
different services (CyL, Aragón, Asturias...) the complexity for the user increases.

```{r option_2_example, eval=FALSE}
get_aemet_daily_data(dates, stations, api, ...) %>%
  dplyr::bind_rows(get_smc_daily_data(dates, stations, api, ...)) %>%
  dplyr::bind_rows(get_meteoclimatic_daily_data(dates, stations, ...))
```

Also, different functions for different periods means more duplicated code between functions:

```{r option_2_example, eval=FALSE}
get_aemet_daily_data <- function(...) {
  # block of common code for all extraction data functions
  {...}
  # block of common code for all aemet extraction data functions
  {...}
  # block of specific code for aemet daily
  {...}
}

get_aemet_monthly_data <- function(...) {
  # block of common code for all extraction data functions
  {...}
  # block of common code for all aemet extraction data functions
  {...}
  # block of specific code for aemet monthly
  {...}
}

get_smc_daily_data <- function(...) {
  # block of common code for all extraction data functions
  {...}
  # block of common code for all smc extraction data functions
  {...}
  # block of specific code for smc daily
  {...}
}

# ...
```

This increase complexity in maintenance, as any change in the common parts of the code has to be replicated
in all functions. You can use helper functions to generate the common part of the code to avoid this, but
it increases package complexity (too many helpers, more time to find where the change must be done...).

## Return

The returned objects for all services must be compatible between them (meaning, they must be able to bind
smoothly). This means same class of object and same columns.  

### Object class

An spatial tibble (dependency on `sf`). It is the most updated standard right now for spatial points data,
tidy and easy to manage.  
Returned objects all should be in the same coordinates system, common latlon (EPSG 4326)

```{r crs}
sf::st_crs(4326)
```

### Object structure

Proposed columns:

  - `stationID`: station code as provided by the service (with enough services, ID collision is a possibility).
  - `service`: service acronym indicating the station origin (aemet, smc...). This avoid collissions betweeen
  stations IDs
  - meteo variables: common names for meteo variables (`temperature`, `temperature_max`, `temperature_min`...)
  - `geometry`: sf geometry column



The expected return value for all services is an spatial tibble object (dependency on `sf`) with common
columns for all services. Each row should be an observation for each timestep (hourly, daily, monthly, instant)
and stations.

## Expected working examples

```{r station_data_mockup, eval=FALSE}
query_options <- aemet_options(
  api = 'xxxxXXXxxx',
  timescale = 'daily',
  dates = lubridate::as_date(lubridate::as_date('1990-01-01'):lubridate::as_date('2020-12-31'))
)

get_meteo_from('aemet', query_options)
```

# Workflow 2: Info access

## OPTION 1

One function, `get_stations_info_for(service)`.

## Return

The returned objects for all services must be compatible between them (meaning, they must be able to bind
smoothly). This means same class of object and same columns.  

### Object class

An spatial tibble (dependency on `sf`). It is the most updated standard right now for spatial points data,
tidy and easy to manage.  
Returned objects all should be in the same coordinates system, common latlon (EPSG 4326)

```{r crs}
sf::st_crs(4326)
```

### Object structure

Proposed columns:

  - `stationID`: station code as provided by the service (with enough services, ID collision is a possibility).
  - info variables: common names for info variables (`elevation`, `lat`, `long`...)
  - `geometry`: sf geometry column


## Expected working examples

```{r station_info_mockup, eval=FALSE}
get_stations_info_for('aemet')
get_stations_info_for('smc')
```

