---
title: "AEMET service"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AEMET service}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
# use eval = NOT_CRAN in the chunks connecting to API, to avoid errors or warnings in CRAN checks
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  purl = NOT_CRAN
)

# env keyring
withr::local_options(list("keyring_backend" = "env"))
```

```{r setup}
library(meteospain)
library(ggplot2)
library(ggforce)
library(units)
library(sf)
library(keyring)
```

# AEMET service

[AEMET](https://www.aemet.es/en/portada) is the Spanish national meteorologic service, and is the national meteorology
authority providing quality data for public and research use, as well as prediction products and disaster
warning system. `meteospain` only access to the automatic meteorological stations network data.

## AEMET options

### Temporal resolution

`meteospain` offers access to the AEMET API at different temporal resolutions:

  - "current_day", returning the latest 24 hours of measures for all or selected stations.
  - "daily", returning the daily aggregated measures for all or selected stations.

In "daily", a `start_date` (and optionally an `end_date`) arguments must be provided, indicating the period
from which retrieve the data.

### Stations

`meteospain` access the data in the AEMET API collecting all stations. If a character vector of stations codes
is supplied in the `stations` argument, a filter step is done before returning the data to maintain only
the stations supplied.

### AEMET API Key

AEMET API only allow access to the data with a personal API Key. This token must be included in the
`api_key` argument of `aemet_options` function.  
To obtain the API Key, please visit https://opendata.aemet.es/centrodedescargas/inicio and follow the
instructions at "Obtencion de API Key".

  > It is not advisable to use the keys directly in any script shared or publicly available (github...),
  neither store them in plain text files. One option is using the
  [keyring package](https://github.com/r-lib/keyring) for managing and accessing keys:
  
  ```{r aemet_key, eval=FALSE}
  install.packages('keyring')
  library(keyring)
  key_set('aemet') # A prompt asking for the secret (the API Key) will appear.
  ```
  

### Examples

```{r aemet_options, eval=NOT_CRAN, results='hide'}
# current day, all stations
api_options <- aemet_options(
  resolution = 'current_day',
  api_key = key_get('aemet')
)
api_options
```

```{r aemet_options_fake, eval=TRUE, echo=FALSE}
# current day, all stations
fake_api <- aemet_options(
  resolution = 'current_day',
  api_key = 'my_api_key'
)
fake_api
```

```{r aemet_options_2, eval=NOT_CRAN, results='hide'}
# daily, all stations
api_options <- aemet_options(
  resolution = 'daily',
  start_date = as.Date('2020-04-25'), end_date = as.Date('2020-05-25'),
  api_key = key_get('aemet')
)
api_options
```

```{r aemet_options_fake_2, eval=TRUE, echo=FALSE}
# daily, all stations
fake_api <- aemet_options(
  resolution = 'daily',
  start_date = as.Date('2020-04-25'), end_date = as.Date('2020-05-25'),
  api_key = 'my_api_key'
)
fake_api
```

## AEMET stations info

Accessing station metadata for AEMET is simple:

```{r aemet_stations, eval = NOT_CRAN}
get_stations_info_from('aemet', api_options)
```

## AEMET data

```{r aemet_data, eval = NOT_CRAN}
api_options <- aemet_options(
  resolution = 'daily',
  start_date = as.Date('2020-04-25'),
  api_key = key_get('aemet')
)
spain_20200425 <- get_meteo_from('aemet', options = api_options)
spain_20200425
```

Visually:

```{r aemet_data_plot, fig.width=7, fig.height=5, fig.align='center', eval = NOT_CRAN}
spain_20200425 %>%
  units::drop_units() %>%
  ggplot() +
  geom_sf(aes(colour = mean_temperature)) +
  scale_colour_viridis_c()

spain_20200425 %>%
  ggplot() +
  geom_histogram(aes(x = precipitation))
```

