---
title: "MeteoCat service"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MeteoCat service}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
# use eval = NOT_CRAN in the chunks connecting to API, to avoid errors or warnings in CRAN checks
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  purl = NOT_CRAN
)

# env keyring
withr::local_options(list("keyring_backend" = "env"))
```

```{r setup}
library(meteospain)
library(dplyr)
library(ggplot2)
library(ggforce)
library(sf)
library(keyring)
```

# MeteoCat service

[MeteoCat](https://www.meteo.cat/) is the Catalonian meteorologic service. It offers access to different
meteorological data and information, being one of their main missions to curate and circulate data from
meteorological stations. `meteospain` only access to the automatic meteorological stations network data.

## MeteoCat options

### Temporal resolution

`meteospain` offers access to the MeteoCat API at different temporal resolutions:

  - "instant", returning the latest 4 hours of measures for all or selected stations.
  - "hourly", returning all measures (some stations has timesteps of 30 min, others 60 min, others more) for
  all or selected stations.
  - "daily", returning daily aggregates for the month in the date provided, i.e. if '2020-04-10' is provided
  as start_date, all daily values for April 2020 will be returned.
  - "monthly", returning monthly aggregates for the year in the date provided, i.e. if '2020-04-10' is
  provided as start_date, all monthly values for 2020 will be returned.
  - "yearly", returning yearly aggregates for all years available. In this case date provided is ignored.

In "daily" and "monthly", a `start_date` argument must be provided, indicating the date from which retrieve
the data as explained earlier. For more info see `vignette('api_limits', package = 'meteospain')`.

### Stations

`meteospain` access the data in the MeteoCat API collecting all stations. If a character vector of stations codes
is supplied in the `stations` argument, a filter step is done before returning the data to maintain only
the stations supplied.

### MeteoCat API Key

MeteoCat API only allow access to the data with a personal API Key. This token must be included in the
`api_key` argument of `meteocat_options` function.  
To obtain the API Key, please visit https://apidocs.meteocat.gencat.cat/ and follow the instructions there.

  > It is not advisable to use the keys directly in any script shared or publicly available (github...),
  neither store them in plain text files. One option is using the
  [keyring package](https://github.com/r-lib/keyring) for managing and accessing keys:
  
  ```{r meteocat_key, eval=FALSE}
  install.packages('keyring')
  library(keyring)
  key_set('meteocat') # A prompt asking for the secret (the API Key) will appear.
  ```
  

### Examples

```{r meteocat_options, eval=NOT_CRAN, results='hide'}
# current day, all stations
api_options <- meteocat_options(
  resolution = 'instant',
  api_key = key_get('meteocat')
)
api_options
```

```{r meteocat_options_fake, eval=TRUE, echo=FALSE}
# current day, all stations
fake_api <- meteocat_options(
  resolution = 'instant',
  api_key = 'my_api_key'
)
fake_api
```

```{r meteocat_options_2, eval=NOT_CRAN, results='hide'}
# daily, all stations
api_options <- meteocat_options(
  resolution = 'daily',
  start_date = as.Date('2020-04-10'),
  api_key = key_get('meteocat')
)
api_options
```

```{r meteocat_options_fake_2, eval=TRUE, echo=FALSE}
# daily, all stations
fake_api <- meteocat_options(
  resolution = 'daily',
  start_date = as.Date('2020-04-25'),
  api_key = 'my_api_key'
)
fake_api
```

## MeteoCat stations info

Accessing station metadata for MeteoCat is simple:

```{r meteocat_stations, eval = NOT_CRAN}
get_stations_info_from('meteocat', api_options)
```

## MeteoCat data

```{r meteocat_data, eval = NOT_CRAN}
api_options <- meteocat_options(
  resolution = 'monthly',
  start_date = as.Date('2020-04-01'),
  api_key = key_get('meteocat')
)
catalunya_2020 <- get_meteo_from('meteocat', options = api_options)
catalunya_2020
```

Visually:

```{r meteocat_data_plot, fig.width=7, fig.height=7, fig.align='center', eval = NOT_CRAN}
catalunya_2020 %>%
  units::drop_units() %>%
  mutate(month = lubridate::month(timestamp, label = TRUE)) %>%
  ggplot() +
  geom_sf(aes(colour = mean_temperature)) +
  facet_wrap(vars(month), ncol = 4) +
  scale_colour_viridis_c()

catalunya_2020 %>%
  mutate(month = lubridate::month(timestamp, label = TRUE)) %>%
  ggplot() +
  geom_histogram(aes(x = precipitation)) +
  facet_wrap(vars(month), ncol = 4)
```

