---
title: "MeteoGalicia service"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MeteoGalicia service}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
# use eval = NOT_CRAN in the chunks connecting to API, to avoid errors or warnings in CRAN checks
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  purl = NOT_CRAN
)

# env keyring
withr::local_options(list("keyring_backend" = "env"))
```

```{r setup}
library(meteospain)
library(ggplot2)
library(ggforce)
library(sf)
```

# MeteoGalicia service

[MeteoGalicia](https://www.meteogalicia.gal/) service offers the data of the galician automatic meteorological
stations network. This network is supported and assessed by the Xunta de Galicia and the data should be
trustworthy.

## MeteoGalicia options

### Temporal resolution

MeteoGalicia API offers data at different temporal resolutions:

  - "instant", returning the latest measures for all or selected stations.
  - "current_day", returning the latest 24 hours of measures for all or selected stations.
  - "daily", returning the daily aggregated measures for all or selected stations.
  - "monthly", returning the monthly aggregated measures for all or selected stations.

In both, "daily" and "monthly", a `start_date` (and optionally an `end_date`) arguments must be provided,
indicating the period from which retrieve the data.

### Stations

MeteoGalicia API allows specifying stations code to narrow the data retrieved. So, a character vector of
station codes can be supplied in the `stations` argument.

### Examples

```{r meteogalicia_options, eval=NOT_CRAN}
# last measure for all stations
api_options <- meteogalicia_options()
api_options

# current day, only some stations
api_options <- meteogalicia_options(
  resolution = 'current_day',
  stations = c('10157', '14000', '10045')
)
api_options

# daily, all stations
api_options <- meteogalicia_options(
  resolution = 'daily',
  start_date = as.Date('2020-04-25'), end_date = as.Date('2020-05-25')
)
api_options

# monthly, some stations
api_options <- meteogalicia_options(
  resolution = 'monthly',
  start_date = as.Date('2020-04-01'), end_date = as.Date('2020-08-01'),
  stations = c('10157', '14000', '10045')
)
api_options
```

## MeteoGalicia stations info

Accessing station metadata for MeteoGalicia is simple:

```{r meteogalicia_stations, eval = NOT_CRAN}
get_stations_info_from('meteogalicia')
```

## MeteoGalicia data

```{r meteogalicia_data, eval = NOT_CRAN}
api_options <- meteogalicia_options(
  resolution = 'daily',
  start_date = as.Date('2020-04-25')
)
galicia_20200425 <- get_meteo_from('meteogalicia', options = api_options)
galicia_20200425
```

Visually:

```{r meteogalicia_data_plot, fig.width=7, fig.height=5, fig.align='center', eval = NOT_CRAN}
galicia_20200425 %>%
  units::drop_units() %>%
  ggplot() +
  geom_sf(aes(colour = max_temperature)) +
  scale_colour_viridis_c()

galicia_20200425 %>%
  ggplot() +
  geom_histogram(aes(x = precipitation))
```

