---
title: "RIA service"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RIA service}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
# use eval = NOT_CRAN in the chunks connecting to API, to avoid errors or warnings in CRAN checks
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  purl = NOT_CRAN
)

# env keyring
withr::local_options(list("keyring_backend" = "env"))
```

```{r setup}
library(meteospain)
library(ggplot2)
library(ggforce)
library(sf)
```

# Red de Información Agroclimática de Andalucía (RIA) service

[RIA](https://www.juntadeandalucia.es/agriculturaypesca/ifapa/riaweb/web/) service offers the data of the
andalucian automatic meteorological stations network. This network is supported and assessed by the
Junta de Andalucía and the data should be trustworthy.

## RIA options

### Temporal resolution

RIA API offers data at different temporal resolutions:

  - "daily", returning the daily aggregated measures for all or selected stations.
  - "monthly", returning the monthly aggregated measures for all or selected stations.

In both, "daily" and "monthly", a `start_date` (and optionally an `end_date`) arguments must be provided,
indicating the period from which retrieve the data.

### Stations

RIA API needs station codes and province codes to retrieve the data. Sadly, RIA doesn't provide unique
station codes, and the uniqueness comes with the province id and station code together. So, to narrow the
data retrieving to the desired stations they must be provided as a character vector of
"province_id-station_code" values (i.e. "14-2") for the `stations` argument. Calling
`get_stations_info_from('ria', ria_options)` will show the station correct codes in the `station_id` column
to take as reference.

### Examples

```{r ria_options, eval=NOT_CRAN}
# default, daily for yesterday
api_options <- ria_options()
api_options

# daily, only some stations
api_options <- ria_options(
  resolution = 'daily',
  stations = c('14-2', '4-2')
)
api_options

# monthly, some stations
api_options <- ria_options(
  resolution = 'monthly',
  start_date = as.Date('2020-04-01'), end_date = as.Date('2020-08-01'),
  stations = c('14-2', '4-2')
)
api_options
```


## RIA stations info

Accessing station metadata for RIA is simple:

```{r ria_stations, eval = NOT_CRAN}
get_stations_info_from('ria', ria_options)
```

## RIA data

```{r ria_data, eval = NOT_CRAN}
api_options <- ria_options(
  resolution = 'daily',
  start_date = as.Date('2020-04-25')
)
galicia_20200425 <- get_meteo_from('ria', options = api_options)
galicia_20200425
```

Visually:

```{r ria_data_plot, fig.width=7, fig.height=5, fig.align='center', eval = NOT_CRAN}
galicia_20200425 %>%
  units::drop_units() %>%
  ggplot() +
  geom_sf(aes(colour = max_temperature)) +
  scale_colour_viridis_c()

galicia_20200425 %>%
  ggplot() +
  geom_histogram(aes(x = precipitation))
```
